### 2019-03-07 manuscript/introduction.tex
polyploidy in the introduction revised

%B: the key alteration below is an attempt at easing the transitions of the complicated existing logical structure, which intermixes advances with setbacks and problems.
%B: the "Here, we" part should maybe come at the bottom of introduction, and it's a little early for here, in the first paragraph.
Species accumulate around the tree of life at different rates.
A possible explanation for this phenomenon is that species possess traits that cause unequal rates of diversification. %B: new.
The search for influential traits has been accelerated by a dramatic increases in data and an advance in methods, which now make use of far more information from phylogenetic trees \citep{maddison_2007}. 
Despite some setbacks \citep{maddison_2015, rabosky_2015, moore_2016}, advances in analytical methods \citep{fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017} are enabling the discovery of traits that alter the diversification patterns. %B: the following is now orphaned, because it's inclusion merits addition of non-trait-focused methods; "potentially interesting biological processes"
Nevertheless, the aim of identifying the focal trait and investigating its association with rates of speciation and extinction remains complicated. 
This is, in part, because the context in which traits occur can lead to complex interactions with the diversification process (see \citet{caetano_2018, herrera_2018}). 
Trait combinations provide examples of such context, and may be of particular interest, given the rapid accumulation of information about traits. %cite?
Two exceedingly well-studied traits, ploidy and breeding system, are of particular interest. 
Their individual state-dependent contributions to the process of diversification, as well as influence on each others' trait transitions, have long been the focus of speculation and study \citep{stebbins1950}.

%B: IMO: ploidy = trait, di-, tri-, tetra-, hexa-, etc. (including "poly-") = trait states; the uncertainty we were discussing is whether "tri-" < "poly-" or "tri-" = "poly-"
%(sorry about the confusion: I accidentally mis-wrote in the commented-out banter with %E.)
% was [B&E banter] and % R- I think the trait is polyploidy and the state is ploidy (as 2, or +2). I will go with everyone else in the literature and use polyploidy
Whole genome duplication is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}.
Such changes is ploidy have the potential to affect many other phenotypes, and therefore alter a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
The prevalence of ploidy variation has been broadly considered as a salient feature of flowering plants for nearly a century \citep{stebbins1938}. 
Polyploids have long been thought to have lower (and even negative) rates of net diversification than diploids \citet{mayrose_2011, mayrose_2015}. 
A number of recent studies, however, find common paleo-polyploidization, including at the base of many highly diverse angiosperm clades \citep{soltis_2014}. 
%B: explanation for this edit, rm: "Using the genomic evidence  from the 1KP project" : this is the data, not the approach; Ks and tree reconciliation are particularly troubling and weak modes of inference. I am okay with having this point to what others have found, but I would prefer to avoid endorsing what I see as a deeply problematic set of methods.
\citet{landis_2018}, for example, found 106 whole genome duplications within angiosperms that, combined with a state-dependent diversification model, resulted in approximately 60\% of whole genome duplication events potentially enhancing net diversification.
The mere occurrence of such ancestral whole genome duplications points to an possibly important role of genome downsizing and diploidization in the diversification process \citep{soltis_2015diploidization, dodsworth_2015}. 
The presence of a significant and positive net diversification rate for (some) diploids can be  the result of a polyploid lineage that has been diploidized. 
Under this scenario, rather than being a macroevolutionary dead-end, polyploidy may instead be positively associated with the diversification process.

>>>>

%B: the key alteration below is an attempt at easing the transitions of the complicated existing logical structure, which intermixes advances with setbacks and problems.
%B: the "Here, we" part should maybe come at the bottom of introduction, and it's a little early for here, in the first paragraph.
Species accumulate around the tree of life at different rates.
A possible explanation for this phenomenon is that species possess various traits or character states that may differentially affect rates of diversification. %B: new.
Dramatic increases in available phylogenetic and phenotypic data %citation
as well as methodological advances %citation
have greatly accelerated the search for such traits that influence diversification traits. Nevertheless, identifying focal traits associated with rates of speciation and extinction remains as a difficult statistical problem (see \citep{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}) . %B: the following is now orphaned, because it's inclusion merits addition of non-trait-focused methods; "potentially interesting biological processes"
In part, this is a difficult problem because speciation and extinction rarely depend on a single trait; the context  in which traits occurs can lead to complex interactions resulting in a heterogenous speciation and extinction rates (\citep{beaulieu_2016, caetano_2018, herrera_2018}). Consequently, examining the association of only one trait when multiple complex interactions are changing diversification patterns can be misleading.

In flowering plants, both ploidy-level and breeding system are well-studied traits that have been linked to diversification but may also affect each other \citep{stebbins1950}. 
%B: IMO: ploidy = trait, di-, tri-, tetra-, hexa-, etc. (including "poly-") = trait states; the uncertainty we were discussing is whether "tri-" < "poly-" or "tri-" = "poly-"
%(sorry about the confusion: I accidentally mis-wrote in the commented-out banter with %E.)
% was [B&E banter] and % R- I think the trait is polyploidy and the state is ploidy (as 2, or +2). I will go with everyone else in the literature and use polyploidy
Whole genome duplication (WGD) is a remarkably common mutation in plants \citep{husband_2013, zenilferguson_2017}, resulting in many plant species with a variety of ploidy levels. The prevalence of ploidy variation has been broadly considered as a salient feature of flowering plants for nearly a century \citep{stebbins1938}.

WGDs have the potential to affect many phentoypes and impact a variety of evolutionary \citep{ramsey_2002} and ecological processes \citep{sessa_2019}.
Polyploids have long been thought to have lower (and even negative) rates of net diversification rate than diploids \citep{mayrose_2011, mayrose_2015}. 
Recent studies, however, find common and numerous paleo-polyploidizations, including some preceding the emergence of highly diverse plant clades \citep{soltis_2014, landis_2018}, suggesting that WGDS have played an important macroevolutionary role driving innovation and diversification in plants.
%B: explanation for this edit, rm: "Using the genomic evidence  from the 1KP project" : this is the data, not the approach; Ks and tree reconciliation are particularly troubling and weak modes of inference. I am okay with having this point to what others have found, but I would prefer to avoid endorsing what I see as a deeply problematic set of methods.
% R- I agree with your point. Tiley, Barker, and Burleigh just published last week a paper about Ks plots and the mixed normal model adjustments to it that "detect" WGDs. They found that in many cases those models lead to overestimating the number of WGDs plus the other issue is using the presence of WGD and then doing something like BAMM that quantitatively does not link via parameters is problematic.
Evidence of paleo-polyploidizations within the genomes of diploid plants also revels the pervasiveness of diploidization, the return of a polyploid to a diploid state throughout the angiosperm phylogeny \citep{soltis_2015diploidization, dodsworth_2015}. Testing for the presence of diploidization in a diversification context is important because the presence of a significant and positive net diversification rate for diploids can be  the result of a polyploid lineage that has been diploidized. %closing this sentence to wrap how previous results would make sense even when polyploidy is the initiator of diversification
